\chapter{Licenses 许可协议}

\section*{English version}

Most of the contents of the notes is from Professor Zhao's blackboard writing and words he had said. The remainder is the combination of articles, webpages, encyclop\ae{}dias and the editor's comprehension. The whole series of notes is licensed under CC BY-NC-SA 4.0 (Attribution-NonCommercial-ShareAlike 4.0 International). Please conform to the license's rules in the course of use:

You are free to:
\begin{itemize}
    \item Share --- copy and redistribute the material in any medium or format
    \item Adapt --- remix, transform, and build upon the material
    The licensor cannot revoke these freedoms as long as you follow the license terms
\end{itemize}

Under the following terms:
\begin{itemize}
    \item Attribution --- You must give appropriate credit, provide a link to the license, and indicate if changes were made. You may do so in any reasonable manner, but not in any way that suggests the licensor endorses you or your use.
    \item NonCommercial --- You may not use the material for commercial purposes.
    \item ShareAlike --- If you remix, transform, or build upon the material, you must distribute your contributions under the same license as the original.
\end{itemize}

Please note that the above terms are only a brief summary of the license. It is not the license itself, nor does it have legal effects. You should carefully review all the terms of the CC BY-NC-SA 4.0 license before using the work. The full text of the license can be found at \url{https://creativecommons.org/licenses/by-nc-sa/4.0/legalcode}.

\section*{中文版本}

笔记的绝大部分内容来自赵老师的板书和上课的讲话，剩余部分为各种论文、网页、百科全书和编者自己的理解的结合体。整套笔记采用署名--非商业性使用--相同方式共享4.0国际协议，请遵守协议条款使用本笔记。

您可以自由地：
\begin{itemize}
    \item 共享 --- 在任何媒介以任何形式复制、发行本作品
    \item 演绎 --- 修改、转换或以本作品为基础进行创作
\end{itemize}

惟须遵守下列条件：
\begin{itemize}
    \item 署名 --- 您必须给出适当的署名，提供指向本许可协议的链接，同时标明是否（对原始作品）作了修改。您可以用任何合理的方式来署名，但是不得以任何方式暗示许可人为您或您的使用背书。
    \item 非商业性使用 --- 您不得将本作品用于商业目的。
    \item 相同方式共享 --- 如果您再混合、转换或者基于本作品进行创作，您必须基于与原先许可协议相同的许可协议分发您贡献的作品。
\end{itemize}

请注意，上面的条款只是对许可协议的简要总结，不是许可协议，也没有法律效力。您应当在使用授权作品前，仔细审阅真正许可协议的所有条款。条款全文可在\url{https://creativecommons.org/licenses/by-nc-sa/4.0/legalcode}查看。

\section*{Font licenses 字体许可协议}

These fonts used in the book are licensed under the \textbf{SIL Open Font License}: Noto Serif, Noto Serif SC (for Chinese characters) and $\mathrm{Computer \ Modern}$. The license could be seen at \url{https://openfontlicense.org/open-font-license-official-text}.

This book also uses \textit{方正仿宋}\_GBK developed by the Founder Corporation.
