\chapter{Foreword: Following the steps of Professor Zhao}

I wrote the first of this series of notes out of pure interest, and somehow the act has continued down to this day, when it has included 30 individual notes and nearly 200 pages, an amount I could never have imagined when I typed the first letters on September 15. And had I not shared the notes to the class group, I would not have been driven by vanity to continue the series.

Professor Zhao has often expressed his hopes of nurturing elite students for CAS and for the Chinese academic world. In such hopes, he extended the contents of his class and included rich content, whether educative or expansive. I hope that we can live up to his expectations and grow to be talents that our country is in desperate needs of.

Now I sincerely thank the support and encouragement that my classmated have provided to me, and whoever pointed my mistakes during the whole process. Also, I am grateful to Shi Jieheng for providing a basic introduction to virtual displacement. Without you, the quality of this series of notes could not reach where it is now.

Professor Zhao, knowledge extensive and profound, has his words and equations in class at hand. I, however, am a small shrimp tumbling slowly forward in a sea of knowledge, and it is inevitable that mistakes and omissions will occur in this set of notes. I would especially appreciate it if you could report any mistakes or inaccuracies that you encounter in the process of reading the book (through Email: \emph{chenyumeng23@mails.ucas.ac.cn}).

\begin{flushright}
    Chen Yumeng

    January 2, 2024 at Yuquan Road
\end{flushright}
